\sysfull

### 2017-11-14 paper/intro.tex
intro: redrawn coalescing figure and prose expanded

\begin{figure}
	%\includegraphics[width=\columnwidth]{figures/}
    \caption{\label{fig:coalesce}Task decomposition is required for intermittent execution but task transitions introduce overhead.}
\end{figure}

@@

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being \emph{inability to adapt} to changing energy conditions.

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. Figure~\ref{fig:coalesce} illustrates how the decomposition into tasks affects
the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor. Such a large task would fail to complete unless the environment provides
sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

@@

\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition approach, that accepts any static decomposition and opportunistically \emph{coalesces} tasks at runtime when there is sufficient energy. The \emph{task coalescing} is a mechanism where atomic tasks are merged into a single (larger) task while preserving atomicity.

Unfortunately, merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution. A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
run-time cannot. To maintain memory consistency in merged tasks, we propose a
mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.

>>>>

\begin{figure}
	\includegraphics[width=0.85\columnwidth]{figures/coalescing-is}
    \caption{\label{fig:coalesce}\textbf{An execution of three
statically-defined tasks without (top) and with (bottom) dynamic
coalescing.} Task coalescing at runtime reduces time and energy overhead in
task-based intermittent programming models by performing fewer commits and
buffering updates to non-volatile (NV) variables in volatile (V) memory.}
\end{figure}

@@

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state~\cite{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor. Such a large task would fail to complete unless the environment provides
sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

@@

\textbf{Task Coalescing.} To address this challenge we propose a
\emph{dynamic} task decomposition approach, that accepts any static
decomposition and opportunistically coalesces tasks at runtime when
there is sufficient energy. \emph{Task coalescing} is a mechanism where
atomic tasks are merged into a single (larger) task while preserving
atomicity of the merged task. If power fails while a colesced task is
running, execution will restart at the first task in the group.
Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing
changes the earlier execution and highlights the change in memory model
necessary to not break atomicity.

Unfortunately, naively merging two atomic tasks might produce a non-atomic
task, that may leave memory inconsistent after a partial execution. A merge
breaks atomicity when the accesses to non-volatile memory from the
previously atomic merged tasks form a \emph{write-after-read (WAR)}
dependency.
%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
run-time cannot. To maintain memory consistency in merged tasks, we propose a
mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.
